Supplement: Analysis and code for 'Testing Perceptual Accuracy in Charts using Surveys

### 2023-07-13 SDSS-2023/testing-charts-accuracy.qmd
jds format for now

---
title: "Testing Charts: viewer's perceptual accuracy in surveys"
short-title: Testing charts in surveys # running title
author:
  - first: Kiegan                 # first name
    first-init: K.              # initial of first name
    last: Rice                  # last name
    marker: 1               # identifier for the affiliation
    footnote-id: 1              # optional identifier for the footnote
    email: foo@bar.com          # optional email address
  - first: Heike
    first-init: H.
    last: Hofmann
    marker: 2
  - first: Nola
    first-init: N.
    last: duToit
    marker: 1
  - first: Edward
    first-init: E.
    last: Mulrow
    marker: 1
footnote:
  - content: Corresponding author
    type: corresp               # set type for the corresponding author
    id: 1
affiliation:
  - institution: National Opinion Research Center (NORC)
    country: United States
    marker: 1
  - institution: Iowa State University
    department: Statistics
    country: United States
    marker: 2
bibliography: references.bib
abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization. We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts. Our findings provide actionable guidance for data visualization practitioners to employ in their work."
keywords: 'list, keywords, here, 3-6, are, fine'
format:
  html: default
  pdf:
    template: template-jdsart.cls.tex
    include-in-header: 
      - text: |
          \usepackage{amsfonts,amsmath,amssymb,amsthm}
          \usepackage{booktabs}
          \usepackage{graphicx}
          \usepackage{longtable}
          \usepackage{hyperref}
      - file: edits.tex
    cite-method: natbib
    keep-tex: true
    keep-md: true
---

@@

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = T, 
                      fig.path="./figures/", # figures are created on the fly, images are pictures like the stimuli
                      fig.env="figure*", 
                      fig.align = "center",
                      out.width = "100%",
                      fig.height = 5,
                      fig.width = 5,
                      dev = "png", dev.args = list(type = "cairo-png")
#                      dpi = 72#, # we can set the resolution up to 300 later.
                      #cache = TRUE) 
)

@@

```{r out.width='\\textwidth'}
#| label: fig-aligned-unaligned
#| fig-cap: "The only difference between the two pairs of rectangles A, B and C, D is their alignment, i.e. A and C are of identical size, as are B and D. While the predominant response for the unaligned pair on the left is 'they are of the same size', more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right."
#| fig-subcap:
#| - "**Unaligned bars**"
#| - "**Aligned bars**"
#| layout-nrow: 1
#| fig-align: "center"
knitr::include_graphics("images/round5-modified.PNG")
knitr::include_graphics("images/round3-modified.PNG")
```

@@

```{r}
#| label: fig-tasks-12
#| fig-cap: "The two stacked barcharts every participant got to see. In each barchart, the two marked tiles are to be compared for their size. In both instances, the tile on the right is (very slightly) larger."
#| fig-subcap:
#| - "**Unaligned bars**"
#| - "**Aligned bars**"
#| layout-nrow: 1
#| fig-align: "center"
#| out-width: 100%
#| fig-asp: 1
knitr::include_graphics("images/vertical-ab.PNG")
knitr::include_graphics("images/vertical-cd.PNG")
```

>>>>

---
title: "Testing Charts: viewer's perceptual accuracy in surveys"
short-title: Testing charts in surveys # running title
author:
  - first: Kiegan                 # first name
    first-init: K.              # initial of first name
    last: Rice                  # last name
    marker: 1               # identifier for the affiliation
    footnote-id: 1              # optional identifier for the footnote
    email: foo@bar.com          # optional email address
  - first: Heike
    first-init: H.
    last: Hofmann
    marker: 2
  - first: Nola
    first-init: N.
    last: duToit
    marker: 1
  - first: Edward
    first-init: E.
    last: Mulrow
    marker: 1
footnote:
  - content: Corresponding author
    type: corresp               # set type for the corresponding author
    id: 1
affiliation:
  - institution: National Opinion Research Center (NORC)
    country: United States
    marker: 1
  - institution: Iowa State University
    department: Statistics
    country: United States
    marker: 2
bibliography: references.bib
abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization. We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts. Our findings provide actionable guidance for data visualization practitioners to employ in their work."
keywords: 'list, keywords, here, 3-6, are, fine'
format:
  pdf:
    documentclass: jds
    fig-width: 6
    fig-asp: 0.618
    fig-align: center
    include-in-header: 
      - text: |
          \usepackage{amsfonts,amsmath,amssymb,amsthm}
          \usepackage{booktabs}
          \usepackage{graphicx}
          \usepackage{longtable}
          \usepackage{hyperref}
      - file: edits.tex
    cite-method: natbib
    keep-tex: true
    keep-md: true
---

@@

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = T, 
                      fig.path="./figures/", # figures are created on the fly, images are pictures like the stimuli
                      fig.env="figure*", 
                 #     fig.align = "center",
                      out.width = "\\textwidth",
                #      fig.height = 5,
                 #     fig.width = 5,
                      dev = "png", dev.args = list(type = "cairo-png")
#                      dpi = 72#, # we can set the resolution up to 300 later.
                      #cache = TRUE) 
)

@@

```{r}
#| label: fig-aligned-unaligned
#| fig-cap: "The only difference between the two pairs of rectangles A, B and C, D is their alignment, i.e. A and C are of identical size, as are B and D. While the predominant response for the unaligned pair on the left is 'they are of the same size', more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right."
#| fig-subcap:
#| - "**Unaligned bars**"
#| - "**Aligned bars**"
#| layout-nrow: 1
#| fig-align: "center"
#| out-width: 50%
knitr::include_graphics("images/round5-modified.PNG")
knitr::include_graphics("images/round3-modified.PNG")
```

@@

```{r}
#| label: fig-tasks-12
#| fig-cap: "The two stacked barcharts every participant got to see. In each barchart, the two marked tiles are to be compared for their size. In both instances, the tile on the right is (very slightly) larger."
#| fig-subcap:
#| - "**Unaligned bars**"
#| - "**Aligned bars**"
#| layout-nrow: 1
#| out-width: 100%
knitr::include_graphics("images/vertical-ab.PNG")
knitr::include_graphics("images/vertical-cd.PNG")
```

### 2023-07-13 analysis/analysis.qmd
jds format for now

coefs_12 %>% 
  mutate(
    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))
  ) %>%
  gt() %>%
  tab_header("Estimates for Model 1",
             " accuracy of responses for unaligned (Question 1) and aligned bars (Question 2)") %>%
  cols_label(
    std.error = "SE",
    statistic = "t-statistic",
    p.value = "p-value"
  ) %>%
  fmt_number(
    columns = c(estimate, std.error),
    decimals = 2
  ) %>%
  fmt_number(
    columns = statistic,
    decimals = 1
  ) %>%
  cols_align(
    columns = p.value,
    align = "right"
  )
```

@@

coefs_certainty %>% 
  mutate(
    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))
  ) %>%
  gt() %>%
  tab_header("Estimates for Model 2.3",
             " certainty of responses by task and correctness of response") %>%
  cols_label(
    std.error = "SE",
    statistic = "t-statistic",
    p.value = "p-value"
  ) %>%
  fmt_number(
    columns = c(estimate, std.error),
    decimals = 2
  ) %>%
  fmt_number(
    columns = statistic,
    decimals = 1
  ) %>%
  cols_align(
    columns = p.value,
    align = "right"
  )
```

@@

```{r estimates, eval=FALSE}
coefs_67 %>% 
  mutate(
    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))) %>%
  gt() %>%
  tab_header("Estimates for Model 3.2",
             " accuracy of responses across pairwise evalulations") %>%
  cols_label(
    estimate = "log odds",
    std.error = "SE",
    statistic = "t-statistic",
    p.value = "p-value"
  ) %>%
  fmt_number(
    columns = c(estimate, std.error),
    decimals = 2
  ) %>%
  fmt_number(
    columns = statistic,
    decimals = 1
  ) %>%
  cols_align(
    columns = p.value,
    align = "right"
  )
```

@@

```{r model-time}
time_mod <- svyglm(time_spent~task, design = survey_67)
summary(time_mod)
```

@@

coefs_7 %>% 
  mutate(
    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))) %>%
  gt() %>%
  tab_header("Estimates for Model 2",
             " accuracy of responses across conditions of Round 7") %>%
  cols_label(
    std.error = "SE",
    statistic = "t-statistic",
    p.value = "p-value"
  ) %>%
  fmt_number(
    columns = c(estimate, std.error),
    decimals = 2
  ) %>%
  fmt_number(
    columns = statistic,
    decimals = 1
  ) %>%
  cols_align(
    columns = p.value,
    align = "right"
  )
```

>>>>

coefs_12 %>% 
  mutate(
    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value)),
        term = ifelse(term=="(Intercept)", "Intercept", term)
  ) %>%
  gt() %>%
  tab_header("Estimates for Model 1",
             " accuracy of responses for unaligned (Question 1) and aligned bars (Question 2)") %>%
  cols_label(
    std.error = "SE",
    statistic = "t-statistic",
    p.value = "p-value"
  ) %>%
  fmt_number(
    columns = c(estimate, std.error),
    decimals = 2
  ) %>%
  fmt_number(
    columns = statistic,
    decimals = 1
  ) %>%
  cols_align(
    columns = p.value,
    align = "right"
  )
```

@@

coefs_certainty %>% 
  mutate(
    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value)),
        term = ifelse(term=="(Intercept)", "Intercept", term)
  ) %>%
  gt() %>%
  tab_header("Estimates for Model 2.3",
             " certainty of responses by task and correctness of response") %>%
  cols_label(
    std.error = "SE",
    statistic = "t-statistic",
    p.value = "p-value"
  ) %>%
  fmt_number(
    columns = c(estimate, std.error),
    decimals = 2
  ) %>%
  fmt_number(
    columns = statistic,
    decimals = 1
  ) %>%
  cols_align(
    columns = p.value,
    align = "right"
  )
```

@@

```{r estimates, eval=FALSE}
coefs_67 %>% 
  mutate(
    term = ifelse(term=="(Intercept)", "Intercept", term),
    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))) %>%
  gt() %>%
  tab_header("Estimates for Model 3.2",
             " accuracy of responses across pairwise evalulations") %>%
  cols_label(
    estimate = "log odds",
    std.error = "SE",
    statistic = "t-statistic",
    p.value = "p-value"
  ) %>%
  fmt_number(
    columns = c(estimate, std.error),
    decimals = 2
  ) %>%
  fmt_number(
    columns = statistic,
    decimals = 1
  ) %>%
  cols_align(
    columns = p.value,
    align = "right"
  )
```

@@

```{r model-time}
time_mod <- svyglm(time_spent~task, design = survey_67)
coef_time <- broom::tidy(time_mod)

coef_time %>% mutate(
    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value)),
    term = ifelse(term=="(Intercept)", "Intercept", term)
  ) %>%
  gt() %>%
  tab_header("Estimates for Model 3.2",
             " time spent (in seconds) on assessing the six pairs") %>%
  cols_label(
    std.error = "SE",
    statistic = "t-statistic",
    p.value = "p-value"
  ) %>%
  fmt_number(
    columns = c(std.error),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(estimate, statistic),
    decimals = 1
  ) %>%
  cols_align(
    columns = p.value,
    align = "right"
  )

@@

coefs_7 %>% 
  mutate(
    term = ifelse(term=="(Intercept)", "Intercept", term),
    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))) %>%
  gt() %>%
  tab_header("Estimates for Model 2",
             " accuracy of responses across conditions of Round 7") %>%
  cols_label(
    std.error = "SE",
    statistic = "t-statistic",
    p.value = "p-value"
  ) %>%
  fmt_number(
    columns = c(estimate, std.error),
    decimals = 2
  ) %>%
  fmt_number(
    columns = statistic,
    decimals = 1
  ) %>%
  cols_align(
    columns = p.value,
    align = "right"
  )
```
